---
title: "Linearregression_challenge"
author: "Temitope Folorunso & Nneka Iduu"
date: "2025-04-02"
output: 
  pdf_document:
    toc: true 
  md_document:
    variant: gfm
---
# Linear Model – 25 pts
## Question 1

1.	4 pts. Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor

```{r setup, include=TRUE}
# Read in data and set variable as a factor
data <- read.csv("PlantEmergence.csv",na="na")
head(data)
# set variable as a factor
data$Treatment <- as.factor(data$Treatment) 
data$DaysAfterPlanting <- as.factor(data$DaysAfterPlanting) 
data$Rep <- as.factor(data$Rep) 
# Load necessary libraries
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)
```
## Question 2

2.	5 pts. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. 

```{r, include=TRUE}
lm_model <- lm(Emergence ~ Treatment*DaysAfterPlanting, data = data)
summary(lm_model)
anova(lm_model)
```
### Question 3

3.	5 pts. Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

```{r, include=TRUE}
lm_model2 <- lm(Emergence ~ Treatment+ DaysAfterPlanting, data = data)
summary(lm_model2)
anova(lm_model2)
```
### Result interpretation



### Question 4

4.	5 pts. Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.
```{r}
lsmeans <- emmeans(lm_model2, ~Treatment) # estimate lsmeans of variety within Treatment
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE) # contrast with Tukey ajustment
Results_lsmeans
```

### Result interpretation


### Question 5

5.	4 pts. The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 

```{r,include=TRUE}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}

plot_cldbars_onefactor(lm_model2,"Treatment")
 
```
### Question 6

6.	2 pts. Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the .md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to your github.

