---
title: "Linear Regression_Coding Challenge 7"
author: "Temitope Ruth Folorunso & Nneka Vivian Iduu"
date: "2025-04-03"
output:
  pdf_document:
    toc: true
  md_document:
    variant: gfm
---

# Linear Model – 25 pts

## Question 1

1.	4 pts. Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor

```{r setup, include=TRUE}
# Read in data and set variable as a factor
data <- read.csv("PlantEmergence.csv",na="na")
str(data)

# Load necessary libraries
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)

# Turn the Treatment , DaysAfterPlanting  and Rep into factors
data <- data %>%
  mutate(
    Treatment = as.factor(Treatment),
    DaysAfterPlanting = as.factor(DaysAfterPlanting),
    Rep = as.factor(Rep)
  )
str(data)

```
## Question 2

2.	5 pts. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. 

```{r, include=TRUE}
lm_model <- lm(Emergence ~ Treatment*DaysAfterPlanting, data = data)
summary(lm_model)
anova(lm_model)
```

## Question 3

3a.	5 pts. Based on the results of the linear model in question 2, do you need to fit the interaction term?

Based on the ANOVA results from Question 2, the interaction term (Treatment:DaysAfterPlanting) is not significant (p = 1), indicating that the effect of Treatment on Emergence does not significantly vary across different DaysAfterPlanting. Therefore, we can simplify the model by removing the interaction term and refitting it with only the main effects.

3b. Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

```{r, include=TRUE}
lm_model2 <- lm(Emergence ~ Treatment+ DaysAfterPlanting, data = data)
summary(lm_model2)
anova(lm_model2)
```
### Result interpretation
The intercept of 182.163 represents the expected Emergence value when the treatment is at its reference level, which is Treatment 1, and the number of days after planting is at the reference level, which is 7 days. This means that when Treatment = 1 and DaysAfterPlanting = 7, the model predicts an Emergence value of 182.163. This is the baseline value for the response variable when both factors are at their reference levels.

The coefficient for Treatment 2 (-134.531) means that, compared to Treatment 1, Emergence is estimated to decrease by 134.531 units when Treatment 2 is applied, with the observation taken 7 days after planting. This difference is statistically significant, as indicated by the p-value (< 2e-16).


## Question 4

4.	5 pts. Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.

```{r}
lsmeans <- emmeans(lm_model2, ~Treatment) # estimate lsmeans of variety within Treatment
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE) # contrast with Tukey adjustment
Results_lsmeans
```

### Result interpretation
The LSMeans show that Treatment 2 has a significantly lower emergence (55.6) compared to all other treatments, which have similar means around 199-200. The compact letter display groups the treatments into two categories: Treatments 5, 3, 6, and 1 are not significantly different from each other, while Treatments 7, 2, and 8 differ significantly. Tukey’s pairwise comparisons confirm that Treatment 2 is significantly lower than all other treatments, while differences within other treatments are generally not significant.


## Question 5

5.	4 pts. The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 

```{r,include=TRUE}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}

plot_cldbars_onefactor(lm_model2,"Treatment")
 
```

### Result Interpretation - significance of letters
The letters help us know which treatments are significantly different from the others. Treatments with the same letter are not significantly different from each other, while treatments with different letters are significantly different.

In the above result, Treatments 1, 4, 8, and 9 share the "ab" label, indicating that they are not significantly different from each other. Treatment 2, marked with a "c," is significantly different from all other treatments. Treatments 3, 5, and 6, labeled with "a," are significantly different from Treatments 2 and 7, but not from each other. Treatment 7, marked with a "b," is only significantly different from Treatments 1, 4, 8, and 9, and not from the others within its group.


## Question 6

6.	2 pts. Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the .md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to your github.

[Link to github](https://github.com/NVI0001/Reproducibility_Project/tree/main/Coding_Challenge7)

