---
title: "Coding_Challenge6"
author: "Temitope Folorunso"
date: "2025-03-22"
output:
  md_document:
    variant: gfm
  pdf_document:
    toc: true
    word_doucment:
---
## Iterations and Functions
Install necessary packages
```{r setup, include=FALSE}
library(ggplot2)
if (!require(drc)) install.packages("drc", repos = "https://cloud.r-project.org")
library(drc)
library(tidyverse)
```

### Making a function in R for example converting Fahrenheit to Celsius

```markdown
### Anatomy of a function

sample.function <- function(... variable goes here ...){
  .... code goes here.... 
  return(... output ...)
}

```

```{r}
#converting Fahrenheit to celsius formular can be re wriiten like this
#(5*(degree_f - 32)/9)
#Convert 32 ºF to celsius
(5*(32 - 32)/9)
#Convert 80 ºF to celsius
(5*(80 - 32)/9)

# making it a function (fahrenheit_temp) is the object i want to input into the function as a value
F_to_C <- function(fahrenheit_temp){
  celsius <- (5*(fahrenheit_temp - 32)/9)
  return(celsius) #output of the function
}  # Function will be saved

# these do the same thing F_to_C is the name of the variable
F_to_C(32)
F_to_C(67)
```

### convert celsius to fahrenheiht function
```{r}
C_to_F <- function(celsius_temp) {
  Farenheiht <- (celsius_temp*(9/5)+32)
  return(Farenheiht)
}

#sample
C_to_F(0)
```

### Iterations using following functions
```markdown
1. rep() 
2. seq() & seq_along() 
3. for loops map() *nested map()
```
```{r}
# Uisng rep() command
rep("A", 3)  # repeats A three times 

rep(c("A", "B"), 5) # repeats A and B, times i.e multiple thins together

rep(c(1,2,3,4), times = 4) # repeats 1,2,3,4, 4 times 

rep(c(1,2,5,3), times = 4, each = 4) # repeats 1 four times, 2 four times, 5 four times, and 2 four times.

# seq() command allows you to write sequences of numbers easily
1:7
seq(1:7) # sequence of numbers 1 to 7
seq(from = 1, to = 7) # sequence of numbers 1 to 7
seq(from = 0, to = 10, by = 2) # sequence of numbers from 0 to 10 by 2s 
# combined seq() and rep()
rep(seq(from = 0, to = 10, by = 2), times = 3, each = 2)

LETTERS
seq_along(LETTERS) #allows to go letter by letter
seq_along(LETTERS[1:5]) # will return 1,2,3,4,5 not the actual letters.
```
### The for loop 
```{r}
# the classic one , it set up iterations i is the iterations or any letters 
for (i in 1:10) {
print(i*2)  
}

# More complicated for loop
for (i in -30:100){
  result <- F_to_C(i)
  print(result)
}
for (i in 0:30){
  result <- C_to_F(i)
  print(result)
}

# printing the results into an object 
celcius.df <- NULL 
for (i in -30:100){   
result_i <- data.frame(F_to_C(i), i)
celcius.df <- rbind.data.frame(celcius.df, result_i) 
}
```

### Complex challenge on how to loop through for non-linear regression
```{r}
# read in the data
EC50.data <- read.csv("Data_Folder/EC50.all.csv", na="na")
head(EC50.data)
isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == "ILSO_5-41c"] ~ 
        EC50.data$conc[EC50.data$is == "ILSO_5-41c"], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), #LL.4 is the name of model
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
    # outputs the summary of the paramters including the estimate, standard
    # error, t-value, and p-value outputs it into a data frame called
    # summary.mef.fit for 'summary of fit'
    summary.fit <- data.frame(summary(isolate1)[[3]]) 
    # outputs the summary of just the EC50 data including the estimate, standard
    # error, upper and lower bounds of the 95% confidence intervals around the
    # EC50
    EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]
#the result above calculate for just one isolate
```
### Creating loop to iterate each isolate
```{r}
#first create a vector
nm <- unique(EC50.data$is)
EC50.ll4 <- NULL # create a null object
for (i in seq_along(nm)) {
  isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == nm[[i]]] ~ 
        EC50.data$conc[EC50.data$is == nm[[i]]], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
    # outputs the summary of the paramters including the estimate, standard
    # error, t-value, and p-value outputs it into a data frame called
    # summary.mef.fit for 'summary of fit'
    summary.fit <- data.frame(summary(isolate1)[[3]])
    # outputs the summary of just the EC50 data including the estimate, standard
    # error, upper and lower bounds of the 95% confidence intervals around the
    # EC50
    EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]
    EC50
    isolate.ec_i <- data.frame(nm[[i]], EC50) # create a one row dataframe containing just the isolate name and the EC50
    colnames(isolate.ec_i) <- c("Isolate", "EC50") # change the column names
    
    # Then we need to append our one row dataframe to our null dataframe we created before
    # and save it as EC50.ll4. 
    EC50.ll4 <- rbind.data.frame(EC50.ll4, isolate.ec_i)
}
#this calculate for all the isolates
#Produce map of the data
ggplot(EC50.ll4, aes(x = EC50)) + geom_histogram() + theme_classic()
```
### Doing the same, using map with tidy verse
```{r}
EC50.data %>%
  group_by(is)%>%
  nest()%>% # allows you to make subdataframe within a dataframe , it keep data tidy in a long format
  mutate(ll.4.mod = map(data, ~drm(.$relgrowth ~ .$conc, 
                              fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                         names = c("Slope", "Lower", "Upper", "EC50")))))%>% #create a new column of summary data of each non-linear regression
  mutate(ec50 = map(ll.4.mod, ~ED(.,  #map is a for loop contain in a function 
                              respLev = c(50), 
                              type = "relative",
                              interval = "delta")[[1]])) %>%
  unnest(ec50) #to get the data out
```


