---
title: "Data wrangling"
author: "Temitope Folorunso"
date: "2025-03-16"
output:
  md_document:
    variant: gfm
  pdf_document:
    toc: true
    word_doucment:
---
```{r Install Tinytex, eval=FALSE}
install.packages('tinytex')
tinytex::install_tinytex()
```


```{r}
library(tidyverse)
```

```{r}
getwd()
microbiome.fungi <- read.csv("Data_Folder/Bull_richness.csv")
str(microbiome.fungi)
```

## Data Exploration
### Select Function select() to select the column you are interested in your data.

```{r}
microbiome.fungi2 <- select(microbiome.fungi, SampleID, Crop, Compartment:Fungicide, richness)
```
### Subsetting/filtering data (i.e., selecting certain rows)
using filter function filter()
```{r}
head(filter(microbiome.fungi2, Treatment == "Conv."))
# A more complex using &
head(filter(microbiome.fungi2, Treatment == "Conv." & Fungicide == "C")) # & means addition of that data
head(filter(microbiome.fungi2, Sample == "A" | Sample == "B")) # means | sample A or sample B
```
### Creating a new colum 
using mutate function mutate()
```{r}
# Create a new column called logRich
head(mutate(microbiome.fungi2, logRich = log(richness))) # create new column named logRich that has the log of richness
# Creating a new column which combines Crop and Treatment  
head(mutate(microbiome.fungi2, Crop_Treatment = paste(Crop, Treatment)))
```
### Using The pipe %>% function
it combine output of one function into input of another function,by connecting them with
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # subsetting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  head() # displaying the first six rows
```
### summarising data using summarise() function
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # subsetting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich)) # calculating overall mean log richness within the conventionally managed treatment

```
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # subsetting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n))
```
### connnecting to plotting
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n)) %>%
  ggplot(aes(x = Fungicide, y = Mean.rich)) + # adding in a ggplot
  geom_bar(stat="identity") +
  geom_errorbar( aes(x=Fungicide, ymin=Mean.rich-std.err, ymax=Mean.rich+std.err), width=0.4) +
  theme_minimal() +
  xlab("") +
  ylab("Log Richness") +
  facet_wrap(~Treatment)
```
### using Joining function
```{r}
# selecting just the richness and sample ID
richness <- microbiome.fungi %>%
  select(SampleID, richness)

# selecting columns that don't include the richness
metadata <- microbiome.fungi %>% 
  select(SampleID, Fungicide, Crop, Compartment, GrowthStage, Treatment, Rep, Sample)

head(metadata)
head(richness)

head(left_join(metadata, richness, by = "SampleID")) # adding the richness data to the metadata based on on the common column of sampleID
```
### Pivoting using pivot_longer() and pivot_wider() Function 
 for pivot_wider() 
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  filter(Class == "Sordariomycetes") %>%
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide
pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
  mutate(diff.fungicide = C - F) %>%  # calculate the difference between the means. 
  ggplot(aes(x = Treatment, y = diff.fungicide)) + # Plot it 
  geom_col() +
  theme_minimal() +
  xlab("") +
  ylab("Difference in average species richness") 
```
